Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-06-29 main.tex
Made the matrices with zeros smaller
In the processes of explaining the instance-timestep relation better in the main text (not done) Fixed typos in author's response
Draft of notation changes done Also tried to cut down to 12 pages a bit more where possible (very close now)

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in such signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, respectively, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation, satisfying these cryptographic requirements, before using it to derive a filter based on the Extended Kalman Filter, with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition capturing different communication assumptions and leakages is given, crucial for cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

    % Bounding rectangles
    \draw [gray] (0,6) rectangle (6.5,1);
    \draw [gray] (0,0.75) rectangle (6.5,-4.25);
\end{tikzpicture}
\caption{Required linear combination aggregation steps at instance $t$.}
\label{fig:agg_steps}
\end{figure}
In addition, we note that an alternative approach to the two-step protocol is computing $\sum^{m}_{j=1}(\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)})$ at the navigator, requiring only values $a_{i,j}^{(t)}, j\in\{1,\dots,m\}$ to be sent from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in section \ref{sec:priv_localization}.

Before giving a formal definition for the construction and security of our desired encryption scheme, we make the following assumptions on the capabilities of the navigator and sensors.
\begin{LaTeXdescription}
    \item[Global Navigator Broadcast] We assume that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.
    \item[Consistent Navigator Broadcast] We assume that broadcast information from the navigator is received equally by all sensors. This means the navigator may not send different weights to individual sensors during a single instance $t$.
    \item[Honest-but-Curious Sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning they are assumed to follow the localisation procedure correctly but may store or use any gained sensitive information.
\end{LaTeXdescription}
We justify the global broadcast assumption by noting that any subset of sensors within the range of the navigator can be considered a complete group and treated as the global set for estimation purposes, generalising the method, while the wide-spread use of cheap non-directional antennas supports the assumption of consistent broadcasts. The final assumption refers to the known problem of misbehaving sensors \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}, often requiring additional complicated detection mechanisms, and will not be considered in this work.

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want the above algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances, and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Ciphertext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\subsection{Estimation problem} \label{subsec:est_problem}
The estimation problem we consider, for which we will reformulate communication to the protocol above, is localisation with range-only sensors. In this work, we will focus on the two-dimensional case for simplicity but will derive methods suitable for extension to a three-dimensional equivalent. The state that we wish to estimate must capture the navigator position, $x$ and $y$, and may contain any other components relevant to the system. It is of the form
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & y & \cdots
    \end{bmatrix}^\top\,. \label{eqn:state_definition}
\end{equation}
This state evolves following some known system model, which at timestep $k$ can be written as
\begin{equation}
    \vec{x}_k = \vec{f}_k(\vec{x}_{k-1}, \vec{w}_k)\,, \label{eqn:system_model}
\end{equation}
with noise term $\vec{w}_k$. Measurements of the system follow a measurement model dependent on sensor $i\in\{1,\dots,n\}$, given by 
\begin{equation}
    z_{k,i} = h_i(\vec{x}_k)+v_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with Gaussian measurement noises $v_{k,i} \sim \mathcal{N}(0,r_{k,i})$ and measurement function
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,,
    \end{split}
\end{equation}
where
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top
\end{equation} 
is the location of sensor $i$.

@@

We will consider a subset of rational numbers in terms of a range $M \in \mathbb{N}$ and fractional precision $\phi \in \mathbb{N}$. This contrasts with the common definition in terms of total and fractional bits \cite{oberstarFixedPointRepresentationFractional2007,schulzedarupEncryptedCooperativeControl2019,farokhiSecurePrivateControl2017}, but allows for a direct mapping to integer ranges which are not a power of two. A rational subset $\mathbb{Q}_{M,\phi}$ is then given by
\begin{equation}
    \mathbb{Q}_{M,\phi} = \left\{q \,\middle|\, \phi q \in \mathbb{N} \wedge -\left\lfloor\frac{M}{2}\right\rfloor \leq \phi q < \left\lfloor\frac{M}{2}\right\rfloor \right\}\,,
\end{equation}
and we can quantize any real number $a$ by taking the nearest rational $q \in \mathbb{Q}_{M,\phi}$, that is, $\argmin_{q\in\mathbb{Q}_{M,\phi}} |a-q|$. In this form, mapping rationals $\mathbb{Q}_{M,\phi}$ to an encryption range $\mathbb{Z}_N$ is achieved by choosing $M=N$ and handling negatives by modulo arithmetic. Additionally, we note that the Q number format requires a precision factor $\phi$ to be removed after each encoded multiplication. This is captured by a third parameter $d$; the number of multiplication factors to add or remove from encodings.

@@

\section{Private Linear Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting the notion of LCAO defined in section \ref{subsec:crypto_problem}. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] Encrypted linear combination at instance $t$ is computed as 
    \begin{equation}
        l^{(t)}_i = H(t)^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(t)}_j)^{a^{(t)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
             l^{(t)}_i &= H(t)^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(t)}_j)^{a^{(t)}_{i,j}} \pmod{N^2} \\
            &= H(t)^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(t)}_{i,j}\omega^{(t)}_j) \pmod{N^2} \\
            &= H(t)^{sk_i}\prod^{m}_{j=1}(N+1)^{a^{(t)}_{i,j}\omega^{(t)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j} \rho_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_i,\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$. Here, $\rho_i^N$ and $H(t)^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] Aggregation is computed as $l^{(t)} = \prod^n_{i=1}l^{(t)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(t)}_{i,j}\omega^{(t)}_j =\\
            &\frac{L((l^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        l^{(t)} &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}a_{i,j}\omega_j}\rho_i^N \pmod{N^2} \\
        \begin{split}
            &= H(t)^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho'^N \pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(t)}_j$ and values $a^{(t)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

\begin{remark}
    Given the construction of the scheme above, it can be seen that any weights $\omega^{(t)}_j$, whose values are known at each sensor, do not need to be broadcast by the navigator. In this case, sensors can replace
    \begin{equation}
        \mathcal{E}_{pk_0}(\omega^{(t)}_j)^{a^{(t)}_{i,j}} = (N+1)^{\omega^{(t)}_ja^{(t)}_{i,j}}\rho_j^N \pmod{N^2}
    \end{equation}
    in \eqref{eqn:our_scheme_lin_comb}, by
    \begin{equation}
        (N+1)^{\omega^{(t)}_ja^{(t)}_{i,j}} \pmod{N^2}\,.
    \end{equation}
    This is due to the removal of $\rho_j^N$ terms during decryption and can be used to reduce the navigator's broadcast communication cost by the number of weights $\omega^{(t)}_j$ that do not hold any information private to the navigator and are known by the sensors in advance.
\end{remark}

>>>>

Typical cryptographic secrecy involves hiding all transferred data such that external parties in the communication network learn no new information from acquired encryptions. This can be achieved with common symmetric and public-key encryption scheme such as AES \cite{gueronIntelAdvancedEncryption2010} and RSA \cite{rivestMethodObtainingDigital1978}, respectively. In some cases however, partial data leakage or encrypted data processing is required for achieving a desired goal which has led to several homomorphic and functional encryption schemes \cite{paillierPublicKeyCryptosystemsBased1999,shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013,chotardDecentralizedMultiClientFunctional2018} finding uses in such signal processing or localisation tasks. In \cite{alanwarPrOLocResilientLocalization2017}, homomorphic encryption is used to make time-independent model-free location estimates where an estimator does not learn sensor measurements or locations. In \cite{aristovEncryptedMultisensorInformation2018}, similar secrecy is achieved with a linear Kalman filter when a hierarchical sensor network is present. In \cite{shiPrivacyPreservingAggregationTimeSeries2011,joyeScalableSchemePrivacyPreserving2013}, privacy-preserving aggregation schemes are presented as a means to compute total powergrid usage without disclosing individual contributions, while in \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, centralised and hidden weighted sum aggregations, pWSAc and pWSAh, are introduced as a means for computing local control inputs in a distributed environment without learning individual inputs.

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation and an implementation that satisfies these requirements, before using it to derive a filter based on the Extended Kalman Filter with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition with different communication assumptions and leakages is given, crucial for its formal security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

@@

    % Bounding rectangles
    \draw [gray] (0,6) rectangle (6.5,1);
    \draw [gray] (0,0.75) rectangle (6.5,-4.25);
\end{tikzpicture}
\caption{Required linear combination aggregation steps at instance $t$.}
\vspace{-\baselineskip}
\label{fig:agg_steps}
\end{figure}
In addition, we note that an alternative approach to the two-step protocol is computing $\sum^{m}_{j=1}(\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)})$ at the navigator, requiring only values $a_{i,j}^{(t)}, j\in\{1,\dots,m\}$ to be sent from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in section \ref{sec:priv_localization}.

Before giving a formal definition for the construction and security of our desired encryption scheme, we make the following assumptions on the capabilities of the participants.
\begin{LaTeXdescription}
    \item[Global Navigator Broadcast] We assume that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.
    \item[Consistent Navigator Broadcast] We assume that broadcast information from the navigator is received equally by all sensors. This means the navigator may not send different weights to individual sensors during a single instance $t$.
    \item[Honest-but-Curious Sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning that they follow the localisation procedure correctly but may store or use any gained sensitive information.
\end{LaTeXdescription}
We justify the global broadcast assumption by noting that any subset of sensors within the range of the navigator can be considered a complete group and treated as the global set for estimation purposes, generalising the method, while the wide-spread use of cheap non-directional antennas supports the assumption of consistent broadcasts. The final assumption refers to the known problem of misbehaving sensors \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}, often requiring additional complicated detection mechanisms, and will not be considered in this work.

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances, and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Ciphertext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\subsection{Estimation problem} \label{subsec:est_problem}
The estimation problem we consider, for which we will reformulate communication to the protocol above, is localisation with range-only sensors. In this work, we will focus on the two-dimensional case for simplicity but will derive methods suitable for extension to a three-dimensional equivalent. The state that we wish to estimate must capture the navigator position, $x$ and $y$, and may contain any other components relevant to the system. It is of the form
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & y & \cdots
    \end{bmatrix}^\top\,. \label{eqn:state_definition}
\end{equation}
This state evolves following some known system model, which at timestep $k$ can be written as
\begin{equation}
    \vec{x}_k = \vec{f}_k(\vec{x}_{k-1}, \vec{w}_k)\,, \label{eqn:system_model}
\end{equation}
with noise term $\vec{w}_k$. Measurements of $\vec{x}_k$ follow a measurement model dependent on sensor $i\in\{1,\dots,n\}$, given by 
\begin{equation}
    z_{k,i} = h_i(\vec{x}_k)+v_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with Gaussian measurement noises $v_{k,i} \sim \mathcal{N}(0,r_{k,i})$ and measurement function
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,,
    \end{split}
\end{equation}
where
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top
\end{equation} 
is the location of sensor $i$.

@@

We will consider a subset of rational numbers in terms of a range $M \in \mathbb{N}$ and fractional precision $\phi \in \mathbb{N}$. This contrasts with the common definition in terms of total and fractional bits \cite{oberstarFixedPointRepresentationFractional2007,schulzedarupEncryptedCooperativeControl2019,farokhiSecurePrivateControl2017}, but allows for a direct mapping to integer ranges which are not a power of two. A rational subset $\mathbb{Q}_{M,\phi}$ is then given by
\begin{equation}
    \mathbb{Q}_{M,\phi} = \left\{q \,\middle|\, \phi q \in \mathbb{N} \wedge -\left\lfloor\frac{M}{2}\right\rfloor \leq \phi q < \left\lfloor\frac{M}{2}\right\rfloor \right\}\,,
\end{equation}
and we can quantize any real number $a$ by taking the nearest rational $q \in \mathbb{Q}_{M,\phi}$, that is, $\argmin_{q\in\mathbb{Q}_{M,\phi}} |a-q|$. In this form, mapping rationals $\mathbb{Q}_{M,\phi}$ to an encryption range $\mathbb{Z}_N$ is achieved by choosing $M=N$ and handling negatives by modulo arithmetic. Additionally, we note that the Q number format requires a precision factor $\phi$ to be removed after each encoded multiplication. This is captured by a third parameter $d$; the number of precision factors present in encodings.

@@

\section{Private Linear Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting the notion of LCAO defined in section \ref{subsec:crypto_problem}. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution.

As aggregation is typically performed on scalar inputs, we extend our notation for the context of multidimensional estimation data by letting an instance $t_{k,\tau}$ uniquely capture the scalar aggregation during an estimation timestep $k$ for a single element with position index $\tau$. To achieve this in practice, any injective function can be used, such as the concatenation $t_{k,\tau}=k\mathbin\|\tau$. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,i\in\{1,\dots,n-1\}$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

    \item[$\mathsf{CombEnc}(t_{k,\tau}, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_j^{(k,\tau)}), a^{(k,\tau)}_{i,j}),\ j\in\{1\dots m\}$] Encrypted linear combination at instance $t_{k,\tau}$ is computed as 
    \begin{equation}
        l^{(k,\tau)}_i = H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
            l^{(k,\tau)}_i &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j) \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad \prod^{m}_{j=1}(N+1)^{a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_i,\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$. Here, $\rho_i^N$ and $H(t_{k,\tau})^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t_{k,\tau}, pk_0, sk_0, l^{(k,\tau)}_1,\shrtdots,l^{(k,\tau)}_n)$] Aggregation is computed as $l^{(k,\tau)} = \prod^n_{i=1}l^{(k,\tau)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j =\\
            &\frac{L((l^{(k,\tau)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        \begin{split}
        l^{(k,\tau)} &= \prod^n_{i=1}H(t_{k,\tau})^{sk_i}\cdot \\
        &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        \begin{split}
            &= H(t_{k,\tau})^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho'^N \pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(k,\tau)}_j$ and values $a^{(k,\tau)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

\begin{remark}
    Given the construction of the scheme above, it can be seen that any weights $\omega^{(k,\tau)}_j$, whose values are known at each sensor, do not need to be broadcast by the navigator. In this case, sensors can replace
    \begin{equation}
        \mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} = (N+1)^{\omega^{(k,\tau)}_ja^{(k,\tau)}_{i,j}}\rho_j^N \pmod{N^2}
    \end{equation}
    in \eqref{eqn:our_scheme_lin_comb}, by
    \begin{equation}
        (N+1)^{\omega^{(k,\tau)}_ja^{(k,\tau)}_{i,j}} \pmod{N^2}\,.
    \end{equation}
    This is due to the removal of $\rho_j^N$ terms during decryption and can be used to reduce the navigator's broadcast communication cost by the number of weights $\omega^{(k,\tau)}_j$ that do not hold any information private to the navigator and are known by the sensors in advance.
\end{remark}
